---
title: "Pre-Interview exercise CGPS"
author: "Rutuja"
date: "2/18/2022"
output: html_document
---


QUAST stands for QUality ASsessment Tool. The tool evaluates genome assemblies by computing various metrics. 
QUAST—a quality assessment tool for evaluating and comparing genome assemblies. QUAST can evaluate assemblies both with a reference genome, as well as without a reference. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
```

```{r message=FALSE, warning=FALSE}
setwd("C:/Users/Rutuja Chhajed/OneDrive - Nexus365/CGPS")
data <- read_csv("data.csv", trim_ws = FALSE)

```

Question 1: How many samples that have failed the contamination check (confindr.contam_status.check_result) and have contamination of > 5% (confindr.percentage_contamination.metric_value)

- 57 out of 1000 samples have failed the contamination check and have >5% contamination. 
-Samples with 0% contamination % pass, and those >0 failed (67)
```{r}

table(data$confindr.contam_status.check_result)
d1<-subset(data, data$confindr.contam_status.check_result=="FAILURE"&data$confindr.percentage_contamination.metric_value>5)
length(unique(d1$sample_name))



ggplot(data, aes(x=factor(confindr.contam_status.check_result),y=confindr.percentage_contamination.metric_value)) + 
  geom_boxplot()
fail<-subset(data,data$confindr.contam_status.check_result=="FAILURE")
hist(fail$confindr.percentage_contamination.metric_value)
```
Question 2: How many samples that have less than  or equal to 50 contigs and a N50>750,000


```{r}
d2<-subset(data, data$`quast.# contigs (>= 1000 bp).metric_value` >=50 & data$quast.N50.metric_value>=750000)
```

Question 3: Select all numeric columns and rename to remove the .quast prefix and .metric_value suffix

Rename confindr.percentage_contamination to contamination_percent


```{r}
library(stringr)
library(dplyr)

nums <- unlist(lapply(data, is.numeric))  

colnames(data)[nums]<-gsub("quast.","",colnames(data[,nums]), ignore.case = TRUE)
colnames(data)[nums]<-gsub(".metric_value","",colnames(data[,nums]), ignore.case = TRUE)
colnames(data)

data <- data %>%
  rename(contamination_percent=confindr.percentage_contamination)

colnames(data)
```



Question4: box plot of Total length (>1000 bp)

Overall, boxplots are extremely helpful in quickly visualization of the central tendency and spread of the data. 
They provide no information on the underlying data distribution and provide a somewhat arbitrary detection of extreme values especially for non-normal data distributions (Kampstra 2008; Krzywinski and Altman 2014).
Outliers:  1.5 times the interquartile range above the upper quartile and below the lower quartile (Q1 - 1.5 * IQR or Q3 + 1.5 * IQR).

```{r}
library(ggplot2)
options(scipen=10000)
p <- ggplot(data, aes(x = factor(1), y = `Total length (>= 1000 bp)`)) + 
  geom_boxplot(width = 0.4, fill = "white",outlier.colour = "red") + labs(x = NULL)  
p
```
Question 5: Convert to 1 observation per row (Tidy format) with only sample) name, Metric and value.

```{r}
col.nums<-unlist(sapply(data[1,],is.numeric))
d3<-pivot_longer(data,which(col.nums==TRUE),names_to = "metric",values_to = "value")
final_data<-d3[,c("sample_name","metric","value")]

d4<-sapply(data, as.character)
d4<-as.data.frame(d4)

d5<-pivot_longer(d4,2:12,names_to = "metric",values_to = "value")
final_data1<-d5[,c("sample_name","metric","value")]

```

Question 6: Violin plots for all numeric variables

Violin plots are similar to box plots, except that they also show the kernel probability density of the data at different values
```{r}
# Basic violin 
final_data$metric<-as.factor(final_data$metric)
p <- ggplot(final_data, aes(x=factor(1), y=as.numeric(value)),fill='white') + 
  geom_violin()+ facet_wrap(~metric,  scales = "free_y")+ylab("Value")+xlab("") +geom_jitter(width=0.15, alpha=0.1,color='red')+theme_bw()

p

```

